CMSC818s: Project 3 \\ A Brief Survey of Modern Virtualization Research

### 2011-12-12 paper.tex
denali summary added

>>>>

Three years after VMWare's appearance, Denali sought to provide a virtualization
solution design with two goals in mind, scalability and security. Denali's authors
envisioned a separation between the management of physical hardware and the management
of software services, transforming how internet services could be deployed. Such a system 
would allow untrusted software to run on third party infrastructure, support the distribution 
of dynamic content in delivery systems, or provide virtual hosting infrastructure.

In order to meet this vision the Denali project introduced \emph{para-virtualization} in the form 
of isolation kernels, adopted complete isolation of guest virtual machines, utilizes lightweight
virtual machines to support over 10,000 VMs on commodity machines. However, Denali diverges from many 
related projects by sacrificing backwards compatibility for its primary objectives, but its authors argue
tha this design trade off is worth the benefits and suggest that legacy operating systems could be
ported to Denali.

At the high level, Denali provides an isolation kernel, a small kernel operating system, running
on top of x86 hardware. The kernel is designed to host many untrusted applications which are isolated
and require little data sharing. Denali multiplexes between the guest operating systems and isolates them
using private virtualized namespaces. Some special privileges, 
like dowloading a remote VM or loading initial disk images, are handled by a supervisor VM.
Since the x86 ISA includes instructioins whose behavior differs
in user and kernel mode, Denali provides a subset of the x86 ISA as well as two virtual instructions.
First, idle-with-timeout lets VMs halt their execution for a bounded amount of physical time or until
an interrupt arrives for the VM in return for a higher priority, 
and secondly Denali provides a virtual instruction which VMs can use to terminate.

Denali's virtual architecture and implementation adhere to a separation between mechanism and policy 
so changes can be easily made. CPU virtualization holds a policy split into two elements, 
a \emph{gatekeeper policy} to choose the subset of active VMs to admit into the system and 
the \emph{scheduler policy} to switch between active VMs. For simplicity, the authors use FIFO
and round robin in the descirbed implementation. 
% maybe new paragraph? and expand the memory virtualization stuff?
Denali's memory virtualization aims to keep a VM's working set in memory, so physical memory 
is redistributed from inactive to active VMs, and swap regions for each VM are used to page its memory.
% maybe new paragraph?
Denali also provides a virtual ethernet NIC for each VM with queues for incoming and outgoing packets.
Incoming packets for each VM are accumulated in the appropriate queues and are later process once the
corresponding VM is given focus by the scheduler, whereas outgoing packets are processed in a round robin manner
from outgoing queues. The fact that Denali's TCP/IP stack runs at the user level has its drawbacks, 
since all packets have to cross the user/kernel boundary. However, the packet dispatch evaluation 
suggested that cost of network virtualization was low and that the the driver was responsible for most of the
costs of packet processing.
% new p?
% The Denali isolation kernel delivers interrupts in batches to VMs
% what else to say about interrupts
% Denali provides simple fixed size virtual disks to VMs and basic devices (keyboard, console, etc) not really important?

The Denali paper provides a comprehensive evaluation of the Denali isolation kernel running
the Ilwaco operating system with basic metrics like switching times between VMs, swap disk
microbenchmarks, and packet dispatch latency, and TCP and HTTP throughput. Some of the most
interesting evaluations were conducted to examine the scalability and performance of Denali.
The authors found that batched interrupts provided a performance gain of 30\% when the number
of VMs exceeded 800, and that using idle-with-timeout in Denali lead to higher performance as
the number of VMs increased.

However, Denali is still not free from bottlenecks, and the authors discuss two major challenges to
Denali's performanc and scalability. First, Denali tries to reduce the amount of per-VM kernel data
that it has too manage by no keeping information for inactive VMs, but it still requires about 81MB
of data for 10,000 virtual machines. The authors suggest that with the growing size of memory,
this bottleneck might be acceptable. Secondly, the paging behavior of the guest is crucial to overall performance,
and while Denali does its best to keep VMs' working sets in memory, a misbehaving guest can severly damage the
performance of the system. One possible solution to this problem might lie with memory management 
and garbage collection, since ensuring good behavior and providing good locality could ameliorate this problem.

% transition to present day?
